How to write a PhD in Biological Sciences: a guide for the uninitiated

### 2021-07-20 2.25-formula.Rmd
Add files via upload
# After copy edits

Very few people can sit down at a desk and write an academic paper from beginning to end. For most of us it takes a lot of hard work over many weeks or even months to write something that can be submitted to peer review. In part two of this book I approach the writing of a chapter or a paper for peer review in the same formulaic way, and will refer to papers or chapters interchangeably.

@@

There is another reason why writing to a formula is recommended. Not only is it easier, but because it is so common and widespread, it also gives the reader a familiarity with reading your work. This means that editors are able to skim through and find exactly what makes your work worth publishing, and it means that those who are only semi-interested are more likely to cite your work.

@@

If we were in my office, at this point in the meeting I usually go to my white board and start drawing funnels (as a metaphor) and blocks. However, others have much better diagrams and here I’ve adapted an idea from Brian McGill’s blog [@mcgill20165]. But you can also find other useful material in other papers [-@turbek2016scientific; @plaxco2010art @tomaska2007teaching]. If you are of the Twitter persuasion, you get lots of useful tips and links from WriteThatPhD and Write4Research.

@@

(ref:formula) **Using a formulaic approach to writing your PhD chapter.** The formula of writing a paper or chapter that we will concentrate on in this book.

@@

### Start with an outline
__So now that you've got the idea that writing to a formula is a good idea, what next? __
Next, I suggest that you begin to plan out your formula as a bullet-point outline. Start by writing a general sentence for the subject of each paragraph. Then use sub-points to plan each sentence within the paragraph. Lastly, annotate these with citations that you want to use. Personally, I'd prefer to see the bare outline, and then a fleshed outline before you start the writing proper. This would allow the advisor and student to get on the same page to be sure that what is planned is thought out.

Example of an outline draft of the [introduction](introduction.html):
This outline is taken from Measey et al [-@measey2020cost]:

__Introduction__

* It is important to assess the impacts of invasive species from published studies in order to compare them
  + Not all studies are equally rigorous
  + Inferences from less robust studies should be treated with caution
* More complex study designs tend to result in greater power to detect effect direction
  + Increasingly complex designs are likely to cost more money to conduct
  + Researchers are likely to conduct cheaper (small scale) experiments before large landscape scale ones
* EICAT and SEICAT rely on studies carried out at scales that are appropriate to the invasive population
  + Cheaper studies will result in impacts of lower magnitude, and/or confidence and not reflect the species’ true impact
  + Poorer regions of the world might be expected to produce impacts of lower magnitude
  + But EICAT is used to inform policy
* There is evidence of inequality in existing impact assessments of birds and amphibians
  + geographic disparity for birds
  + This could come from funding disparities
  + how can we achieve a representative impact level for all alien species, without enough peer reviewed studies?
  + will the impact score of species be independent of research costs?
* Aims
  + Whether studies that result in higher impact scores cost more to conduct
  + Whether impact is related to study complexity or time taken to conduct the study
  + Whether EICAT and SEICAT scores demonstrate changes through time
  + Whether different EICAT and SEICAT scores occur with equivalent confidence

__Materials and Methods__

Paragraphs are indicated by the solid dots, and points to make inside the paragraphs hollow dots.

@@

In the next stage, fleshing out, I ask that citations for statements above be added along with any examples. Below is an example of the citations added to paragraph 4 above.

- There is evidence of inequality in existing impact assessments of birds, amphibians and plants with lots of DD species (30%; Evans et al. 2016; 38%; Kumschick et al. 2017; 79%; Rumlerová et al. 2016)
  + Geographic disparity for birds (Evans and Blackburn 2020)
  + Examples of birds from Evans et al (2020)
  + This could come from funding disparities
  + How can we achieve a representative impact level for all alien species, without enough peer reviewed studies?
  + Will the impact score of species be independent of research costs?

>>>>

Very few people can sit down at a desk and write an academic paper from beginning to end. For most of us, it takes a lot of hard work over many weeks or even months to write something that can be submitted for peer review. In the third part of this book, I approach the writing of a chapter or a paper for peer review in the same formulaic way, and will refer to papers or chapters interchangeably.

@@

There is another reason why writing to a formula is recommended. Not only is it easier, but because it is so common and widespread, it also gives the reader a familiarity with reading your work. This means that editors are able to skim through and find exactly what makes your work worth publishing, and it means that those who are only semi-interested are more likely to find what they need to cite in your work.

@@

If we were in my office, at this point in the meeting I usually go to my white board and start drawing funnels (as a metaphor) and blocks. However, others have much better diagrams and here I’ve adapted an idea from Brian McGill’s blog [-@mcgill20165]. But you can also find other useful material in other papers [@turbek2016scientific; @plaxco2010art @tomaska2007teaching]. If you are of the Twitter persuasion, you get lots of useful tips and links from WriteThatPhD and Write4Research.

@@

(ref:formula) **Using a formulaic approach to writing your PhD chapter.** This diagram represents the formula of writing a paper or chapter that we will concentrate on in this book. Imagine the flow of the reader being like a liquid that passes from one section to the next.

@@

### Start with an outline
__So now that you've got the idea that writing to a formula is a good idea, what next? __
Next, I suggest that you begin to plan out your formula as a bullet-point outline (Figure \@ref(fig:outline)). Start by writing a general sentence for the subject of each paragraph. Then use sub-points to plan each sentence within the paragraph. Lastly, annotate these with citations that you want to use. Personally, I'd prefer to see the bare outline, and then a fleshed outline before you start the writing proper. This allows advisor and student to get on the same page to be sure that what is planned is thought out.

(ref:outline) **Example of an outline draft of the introduction.** In this outline taken from Measey et al. (2020), paragraphs are paragraphs are indicated by the solid dots, and points to make inside the paragraphs by hollow dots. Each paragraph has an argument sketched out within it.

```{r outline, echo=FALSE, out.width = '95%', fig.cap="(ref:outline)"}
 knitr::include_graphics("figures/outline.png")
```

@@

In the next stage, 'fleshing out', I ask that citations for statements above be added along with any examples. Figure \@ref(fig:outline-fleshed) is an example of the citations added to paragraph 4 above.

(ref:outline-fleshed) **Example of a fleshed out outline of a paragraph in the introduction.** In this outline taken from Measey et al. (2020), citations and relevant figures have been inserted for each of the statements.

```{r outline-fleshed, echo=FALSE, out.width = '95%', fig.cap="(ref:outline-fleshed)"}
 knitr::include_graphics("figures/outline-fleshed.png")
```

### 2021-07-20 2.27-data.Rmd
Add files via upload
# After copy edits

## Data Management {#data}

@@

At the time of writing, it is very easy to keep copies of documents on the cloud, or even write straight onto the cloud. Talk to your advisor about what options are available in your institution for backing up your work. Make a plan early on, before you start with your study, and then stick with that backup plan as you move through your PhD. A really good option is to start working in one of the Open Science platforms. For example, OSF allows you to create a project where you can continue to use lots of different cloud based resources, and provides an integrated home for them (for an introductory video see here). OSF is also a great platform as it is free to use, and as long as individual files are less than 5 Gb, OSF will host them. This means that for most projects, raw data files can be stored on the cloud at no cost to the project.

### Should the cloud be the only copy of your data?

No. I suggest that you keep a hard drive with a copy of your data that you regularly deposit on a monthly basis, or after significant bouts of data collection such as field trips or experiments. A hard drive has the advantage that in the case of an internet outage, you always have access to your data.

@@

### Backing up in the field

If you are going to be doing lots of field work, then you might want to make a specific plan about backing up data to the cloud during the field time. It may seem expensive, but I'm sure that it won't be as expensive as losing the data. Alternatively, or in addition, you may want to take a portable storage option.

I still prefer field notes and data to be written into a book (preferably with waterproof paper such as [rite in the rain](https://www.riteintherain.com/)). This allows you to make a photographic backup of the hard copy. Similarly, you can design datasheets that you use in the field with a pencil, and then photograph them to make a hard copy. I don't like entering data directly into a laptop or other device. For me, the advantage is that it is too easy to mis-type a data point into a device, and then there is no record about what it should have been. These days there are multiple different ways to enter data into devices (including dictation), and so you should make your own plan about what you feel most comfortable with. If you do make a hard copy, you will then have to transcribe it later. If you make mistakes, you can go back to the hard copy.

### Make a data management plan, and stick to it

>>>>

# Data Management {#data}

@@

At the time of writing, it is very easy to keep copies of documents in the Cloud, or even write straight into the Cloud. Talk to your advisor about what options are available in your institution for backing up your work. Make a plan early on, before you start with your study, and then stick with that backup plan as you move through your PhD. A really good option is to start working in one of the Open Science platforms. For example, OSF allows you to create a project where you can continue to use lots of different Cloud based resources, and provides an integrated home for them. OSF is also a great platform as it is free to use, and as long as individual files are less than 5 Gb, OSF will host them. This means that for most projects, raw data files can be stored in the Cloud at no cost to the project.

## Should the cloud be the only copy of your data?

No. I suggest that you keep a hard drive with a copy of your data that you regularly update on a monthly basis, or after significant bouts of data collection such as field trips or experiments. A hard drive has the advantage that in the case of an internet outage, you always have access to your data.

@@

## Backing up in the field

If you are going to be doing lots of field work, then you might want to make a specific plan about backing up data to the cloud during field time. It may seem expensive, but I'm sure that it won't be as expensive as losing the data. Alternatively, or in addition, you may want to take a portable storage option.

I still prefer field notes and data to be written into a book (preferably with waterproof paper such as '[rite in the rain](https://www.riteintherain.com/)'). This allows you to make a photographic backup of the hard copy. Similarly, you can design datasheets that you use in the field with a pencil, and then photograph them to make a soft copy from your hard copy. I don't like entering data directly into a laptop or other device. For me, the advantage is that it is too easy to mis-type a data point into a device, and then there is no record about what it should have been. These days there are multiple different ways to enter data into devices (including dictation), and so you should make your own plan about what you feel most comfortable with. If you do make a hard copy, you will then have to transcribe it later. If you make mistakes, you can go back to the hard copy.

## Make a data management plan, and stick to it

### 2021-07-20 3.01-Title.Rmd
Add files via upload
# After copy edits

For the first draft, type away confident in the knowledge that 90% of the first draft will get junked or changed so much that you won’t recognise it. That’s ok, and it’s not a waste of your time. In fact, the first draft is invaluable in getting your ideas and perspective down on paper, allowing you to better organise and sift through your thoughts to bring clarity to your manuscript. Allow yourself a reward for finishing your first draft, something that you won’t do until it’s done. Now make life easy for yourself by changing the order of what you write.

Although I have placed the contents of this chapter as you’d expect to see them in a typical paper, this is not the order I would suggest that you write them in. The most important part of writing up your chapter or paper is to have your hypothesis or question clearly stated. Once you have this, it will help you to know the literature that is already there and so writing the introduction next seems logical. Of course, you may already have [your proposal](proposal.html) to work with, in which case you can start by reviewing what you already have and deciding whether or not you need to tweak this, or if you need a fresh start.

The methods and materials are the easiest section to write first, especially if you've already done the practical work. This section will also help you focus on the analysis that you are doing for the results. Again you should review your proposal and see whether there were any significant changes to your planned methodology and note them down. This will not be an issue even if you have [preregistered your proposal](transparency.html), you simply need to provide a rational explanation for why things changed.

Next write the results. Remember that you are only looking to respond to your hypotheses, preferably using your (preregistered) proposal plans. Tables and figures are compiled along with the results. Now you have the materials and methods together with the results in full. I suggest that you make a list of all the relevant discussion points you want to make and then turn your attention to the introduction. Begin with an outline, before you flesh it out.

@@

Getting authors' names correct, their correct addresses (many people have more than one affiliation), can be tricky. Check that you have all of the information required of a title page before you submit. While you need the names of your co-authors to be displayed correctly, you do not need their titles (no Dr. or Prof., etc.).

Indicate clearly who is the corresponding author. Normally, if you have done all of the rest of the work, the corresponding author should be you. You need to learn how to start this role at some point, so it might as well be now. Having the correct name and address will be important to each author. Along with the author names and addresses you should record the ORCID [see part 3](ORCID.html) of each author. The ORCID is simply a unique identification code for individual researchers. ORCID is a non-profit organisation and there’s nothing sinister in signing up. If they (or you) don’t have one, then you should ask them to create one before the submission. It takes less than 5 minutes. Go to www.orcid.org

Note that the same information given in the title page will also be needed when you enter the information into the journal submission site.

Work closely with your advisor, and make sure that your co-authors are aware of the correspondence that arrives from the journal. Use an address for correspondence that you know you check regularly. It;s also worth checking the spam of your email regularly so that you catch any correspondence that might have been misclassified. The golden rule for the corresponding author is the timely communication with the other authors when decisions have to be made.

## The Title

The title of your paper, chapter or book is the first thing that any reader will read, and so should be well considered. If your reader cannot quickly understand your title, the chances are that they won’t bother reading any further. Your title will be your selling point, and your aim is to use it to draw your readership in. Once you’ve managed to invegle your readers to download your paper, your title is also their hook for remembering your paper in their database of thousands of others. Having their key word in your title will help here, and as ever with writing your challenge is to think like your reader. The best titles are those that sum up the entire study in five to seven words. This is best done in a narrative that tells the story ([see part 2](storytelling.html)) of your manuscript in its entirety. This may sound daunting, but you should get into the habit of summing up your story quickly (for friends, relatives and work colleagues). Then it’s a question of refining this story into the short single sentence that makes up the title. While the narrative approach may not work for you, you do want the title to provide enough information so that the potential reader knows what they will find before they open it. Your title doesn’t just have to work for you, it needs to work for a wide audience.

Some people are excellent at writing titles that contain puns of well known phrases or sayings. These can be brilliant, working both to inform what’s in the paper as well as providing some familiar input that helps retain them in memory. However, many fail to do either and are simply a waste of space. If you are tempted to use a pun as your title, make sure that it is widely appreciated, and not just among your co-authors and lab.

You don’t have to come up with a killer title from day one. Most of my manuscripts have a working title that gets revised as I write, and is always open for change before submission. If you have great ideas for a title, do note them down. I find that the more options I have, the more likely I am to come up with something that works for everyone. It also helps to mix and match from a set of candidate titles. Once you have come up with something that looks good to you and your colleagues, test it by entering it into your database of choice (with the default being Google Scholar - see Part 2). Your first 10 results should include a set of papers that you have likely cited in the upper area of your introduction. If you don’t recognise any out of the top 10, it’s time to look at another of your candidate titles.

### Some title ideas to start you off
- Don’t start by looking for the best title
  + Write a number of candidate titles and ask your co-authors to vote for their favorites.
- The shorter and catchier your title can be the better: 6 words (±1) is an ideal.
  + Allow yourself a longer subtitle if needed, but don’t go over 20 words total (some journals may limit your total to less). 
  + Consider the (former) 120 character limit of a Tweet as an upper limit
- Do include your principal finding if possible
- Include as many key words as you can

### Things to avoid in your title
 - Don’t feel obliged to include [taxonomic terms](taxonomy.html) unless it is relevant or compulsory
   + Some (taxonomically minded) journals will insist on the species name followed by the taxonomic authority, and/or the family and order, in your title
 - Avoid obscure or specialist words that won’t be understood by your readership
   + There are times when key words are necessarily specialised and your readership will expect this, but simpler words in your title will open up your readership which will otherwise remain narrow
 - Don’t simply define the scope of your work without including your content.
 - Don’t have your title as a question. Rather provide the answer!

## Key Words
The key words are a way for readers to find your content with searches. Typically, the advice is to use words that are not in the title or abstract. This is because many databases have combined searching facilities for title, abstract and key words. I struggle to think of appropriate key words. Some of the big idea words and short phrases from the introduction. Then I tend to look at articles in the same genre, and see what key words they have used. As with the title, I suggest that you enter your chosen key words into your literature database of choice [see part 2](database.html) and see what comes back. You should see a group of papers that look wholly familiar and preferably those that are already cited in your manuscript. If not, it’s time to look again.

>>>>

For the first draft, type away confident in the knowledge that 90% of the first draft will get junked or changed so much that you won’t recognise it. That’s ok, and it’s not a waste of your time. In fact, the first draft is invaluable in getting your ideas and perspective down on paper, allowing you to better organise and sift through your thoughts and bring clarity to your manuscript. Allow yourself a reward for finishing your first draft, something that you won’t do until it’s done.

You can make life easier for yourself by changing the order of what you write. Although I have placed the contents of this chapter as you’d expect to see them in a typical paper, this is not the order I would suggest that you write them in. The most important part of writing up your chapter or paper is to have your hypothesis or question clearly stated. Once you have this, it will help you to know the literature that is already there and so writing the introduction next seems logical. Of course, you may already have [your proposal](#proposal) to work with, in which case you can start by reviewing what you already have and deciding whether or not you need to tweak this, or if you need a fresh start.

The methods and materials are the easiest section to write first, especially if you've already done the practical work. This section will also help you focus on the analysis that you are doing for the results. Again you should review your proposal and see whether there were any significant changes to your planned methodology and note them down. This will not be an issue even if you have [preregistered your proposal](#transparency), you simply need to provide a rational explanation for why things changed.

Next, write the results. Remember that you are only looking to respond to your hypotheses, preferably using your (preregistered) proposal plans. Tables and figures are compiled along with the results. Now you have the materials and methods together with the results in full. I suggest that you make a list of all the relevant discussion points you want to make and then turn your attention to the introduction. Begin with an outline, before you flesh it out.

@@

Getting authors' names correct, their correct addresses (many people have more than one affiliation), can be tricky. Check that you have all of the information required for a title page before you submit. While you need the names of your co-authors to be displayed correctly, you do not need their titles (no Dr. or Prof., etc.).

Indicate clearly who is the corresponding author. Normally, if you have done all of the rest of the work, the corresponding author should be you. You need to learn how to start this role at some point, so it might as well be now. Having the correct name and address will be important to each author. Along with the author names and addresses you should record the ORCID of each author. The ORCID is simply a unique identification code for individual researchers. ORCID is a non-profit organisation and there’s nothing sinister in signing up. If they (or you) don’t have one, then you should ask them to create one before the submission. It takes less than 5 minutes. Go to www.orcid.org

## The title

The title of your paper, chapter or book is the first thing that any reader will read, and so should be well considered. If your reader cannot quickly understand your title, the chances are that they won’t bother reading any further. Your title will be your selling point, and your aim is to use it to draw your readership in. Once you’ve managed to inveigle your readers to download your paper, your title is also their hook for remembering your paper in their database of thousands of others. Having their key-word in your title will help here, and as ever with writing your challenge is to think like your reader. The best titles are those that sum up the entire study in five to seven words. This is best done in a narrative that tells the story ([see part 2](#storytelling)) of your manuscript in its entirety. This may sound daunting, but you should get into the habit of summing up your story quickly (for friends, relatives and work colleagues). Then it’s a question of refining this story into the short single sentence that makes up the title. While the narrative approach may not work for you, you do want the title to provide enough information so that the potential reader knows what they will find before they open it. Your title doesn’t just have to work for you, it needs to work for a wide audience.

Some people are excellent at writing titles that contain puns of well-known phrases or sayings. These can be brilliant, working both to inform what’s in the paper as well as providing some familiar input that helps retain them in memory. However, many fail to do either and are simply a waste of space. If you are tempted to use a pun as your title, make sure that it is widely appreciated, and not just among your co-authors and lab.

You don’t have to come up with a killer title from day one. Most of my manuscripts have a working title that gets revised as I write, and is always open for change before submission. If you have great ideas for a title, do note them down. I find that the more options I have, the more likely I am to come up with something that works for everyone. It also helps to mix and match from a set of candidate titles. Once you have come up with something that looks good to you and your colleagues, test it by entering it into your database of choice (with the default being Google Scholar). Your first 10 results should include a set of papers that you have likely cited in the upper area of your introduction. If you don’t recognise any out of the top 10, it’s time to look at another of your candidate titles.

### Some title ideas to start you off
- Don’t start by looking for the best title
  + Write a number of candidate titles and ask your co-authors to vote for their favourites.
- The shorter and catchier your title can be the better: 6 words (±1) is an ideal.
  + Allow yourself a longer subtitle if needed, but don’t go over 20 words total (some journals may limit your total to less). 
  + Consider the (former) 120 character limit of a Tweet as an upper limit
- Do include your principal finding if possible
- Include as many key-words as you can

### Things to avoid in your title
 - Don’t feel obliged to include [taxonomic terms](#taxonomy) unless it is relevant or compulsory
   + Some (taxonomically minded) journals will insist on the species name followed by the taxonomic authority, and/or the family and order, in your title
 - Avoid obscure or specialist words that won’t be understood by your readership
   + There are times when key words are necessarily specialised and your readership will expect this, but simpler words in your title will open up your readership which will otherwise remain narrow
 - Don’t simply define the scope of your work without including your content.
 - Don’t have your title as a question. Rather provide the answer!

## Key-words
The key-words are a way for readers to find your content with searches. Typically, the advice is to use words that are not in the title or abstract. This is because many databases have combined searching facilities for title, abstract and key-words. I struggle to think of appropriate key-words, and so make a list of some of the [big idea](#bigidea) words and short phrases from the introduction. Then I tend to look at articles in the same genre, and see what key-words they have used. As with the title, I suggest that you enter your chosen key-words into your literature database of choice [see part 2](#database) and see what comes back. You should see a group of papers that look wholly familiar and preferably those that are already cited in your manuscript. If not, it’s time to look again.

### 2021-07-20 3.02-Abstract.Rmd
Add files via upload
# After copy edits

## The Abstract {#abstract}

@@

### What is the abstract?
The abstract is a concise paragraph that sums up the major points of your manuscript so that the potential reader will be able to assess whether or not they want to read the entire paper. It is an abstract of the entire document. It should not be abstract!

### So what would a good abstract contain?
A good abstract is a summary of the highlights of the paper. You can’t hope to include all of the results, but you should include sample sizes, variables and relevant statistics when they support your major hypotheses. You must include the broader subject area that your study fits into, and show how your results are relevant to this.

None of this is easy, and you should not expect to write your abstract in a single sitting. It will likely require multiple iterations, and some intense word-smithing to make it as good as it can be. Abstracts almost always have a word limit, and that makes it challenging and means that you have to be concise [see part 2](concise.html).

Increasingly, you’ll hear that a good abstract is citable. This means that it contains enough information that someone knows that it can be used to cite for a specific fact. Of course, these people should download and read the entire study, but there is no guarentee that they will.

### Where do you start?
Just like planning your writing in general, I’d suggest starting your abstract with an outline. Use bullet points to make a list of things that you feel that you should include. Rearrange your list until you have all the introduction points at the start, results in the middle and discussion at the end.

@@

### When do you write your abstract?
Although the abstract comes on page 1 of your manuscript, only try to write it once you’ve got to the end of the process of writing your manuscript so that you are aware of all the major points that you need to include. Maybe once you've [submitted your second draft](goodenough.html) of the manuscript to your advisor (given that you are both happy with the content), you can formulate the abstract.

### Do abstracts for conferences differ?
Yes. It’s likely that an abstract for a talk will not be the same, unless you have already published the study: but even then it’s probably worth re-writing it. I’d suggest that your conference abstract be more descriptive and thought provoking, such that it is aimed toward the specific audience of the conference. Pose questions that you will answer in your talk.

### Where do people go wrong in writing the abstract?
- The most common mistake is getting the balance wrong. I often see an abstract that gets to the results, runs out of space and simply stops. It is important to have a statement about your interpretation of the results which relates back to your hypothesis.
- Another common mistake is to have a very simply worded abstract that conveys very little information. There is definitely going to be a lot to put in, so expect to write something too long with too much content, and then cut it down by discussing with your advisor what can be left out.
- Text from the main document is copied into the abstract. I’ve written about this [elsewhere](plagiarism.html), but you should never copy parts of your writing. You could start like this, but wordsmithing must change the words and their order substantially. Otherwise, when the first sentence of the abstract is the same as the first sentence of the manuscript, it will really turn off your readers.
- Too many statistics or no numbers are also common mistakes. Adding numbers and statistics are important, especially when they take the place of more words. 
- There is only time for one or two comments on the discussion, so pick out something important. Look through your written discussion and rank the points that you make in terms of interest to others. Can you combine two or more points into a single sentence?
- The introduction of the abstract must frame the bigger question, and the discussion must show how your data responds to this. 
- Don’t be tempted to overreach with your claims in the abstract (or the main text). Include major caveats when they are relevant.
- Finish with something conclusive and strong, preferably how this study changes the understanding, and not a caveat or suggested further study.
- The abstract does not contain citations (or very rarely and only if unavoidable), so don’t put any in!
  + If you have to have a citation in your abstract, then it must include sufficient information that it can be found without opening up the paper: _first author, abbreviated journal name, volume, first page_

>>>>

# The Abstract {#abstract}

@@

## What is the abstract?
The abstract is a concise paragraph that sums up the major points of your manuscript so that the potential reader will be able to assess whether or not they want to read the entire paper. It is an abstract of the entire document.

## So what would a good abstract contain?
A good abstract is a summary of the highlights of the paper. You can’t hope to include all of the results, but you should include sample sizes, variables and relevant statistics when they support your major hypotheses. You must include the broader subject area that your study fits into, and show how your results are relevant to this.

None of this is easy, and you should not expect to write your abstract in a single sitting. It will likely require multiple iterations, and some intense word-smithing to make it as good as it can be. Abstracts almost always have a word limit, and that makes it challenging and means that you have to be concise [see part 2](#concise).

Increasingly, you’ll hear that a good abstract is citable. This means that it contains enough information that someone knows that it can be used to cite for a specific fact. Of course, these people should download and read the entire study, but there is no guarantee that they will.

## Where do you start?
Just like planning your writing in general, I’d suggest starting your abstract with an outline. Use bullet points to make a list of things that you feel that you should include. Rearrange your list until you have all the introduction points at the start, results in the middle and discussion at the end.

@@

## When do you write your abstract?
Although the abstract comes on page 1 of your manuscript, only try to write it once you’ve got to the end of the process of writing your manuscript so that you are aware of all the major points that you need to include. Maybe once you've [submitted your second draft](#goodenough) of the manuscript to your advisor (given that you are both happy with the content), you can formulate the abstract.

## Do abstracts for conferences differ?
Yes. It’s likely that an abstract for a talk will not be the same, unless you have already published the study: but even then it’s probably worth re-writing it. I’d suggest that your conference abstract be more descriptive and thought-provoking, such that it is aimed toward the specific audience of the conference. Pose questions that you will answer in your talk.

## Where do people go wrong in writing the abstract?
- The most common mistake is getting the balance wrong. I often see an abstract that gets to the results, runs out of space and simply stops. It is important to have a statement about your interpretation of the results which relates back to your hypothesis.
- Another common mistake is to have a very simply worded abstract that conveys very little information. There is definitely going to be a lot to put in, so expect to write something too long with too much content, and then cut it down by discussing with your advisor what can be left out.
- Text from the main document is copied into the abstract. I’ve written about this [elsewhere](#plagiarism), but you should never duplicate text within your document. You could start like this, but wordsmithing must change the words and their order substantially. Otherwise, when the first sentence of the abstract is the same as the first sentence of the manuscript, it will really turn off your readers.
- Too many statistics or their complete absence are also common mistakes. Adding numbers and statistics are important, especially when they can take the place of more words. 
- There is only time for one or two comments on the discussion, so pick out the most important. Look through your written discussion and rank the points that you make in terms of interest to others. Can you combine two or more points into a single sentence?
- The introduction of the abstract must frame the bigger question, and the discussion must show how your data responds to this. 
- Don’t be tempted to overreach with your claims in the abstract (or the main text). Include major caveats when they are relevant.
- Finish with something conclusive and strong, preferably how this study changes the understanding, and not a caveat or suggested further study.
- The abstract does not contain citations (or very rarely and only if unavoidable), so don’t put any in!
  + If you have to have a citation in your abstract, then it must include sufficient information that it can be found without opening up the paper: _first author, abbreviated journal name, volume, first page_.

### 2021-07-20 3.03-Introduction.Rmd
Add files via upload
# After copy edits

## The Introduction {#introduction}

### Why do we need an introduction?
The introduction and discussion appear to be two common stumbling points for students writing chapters or manuscripts. First, what to put in and what to leave out. And second how to construct it. They are complimentary, and if written well can be read without the Materials and Methods or Results, with the reader still gaining all of the major information of the manuscript.

The introduction is going to be the first part of your manuscript that anyone reads. Yes, they’ve already taken in the title and the abstract. These are almost like bait to draw readers in. The meat starts with the introduction. And it’s not only important to get all the correct content in there, it’s also important not to give misleading information that might distract the reader.

When you write the introduction to a paper you must not be tempted to stray away from what we are aiming to introduce. The aim of our introduction is to explain to the reader the hypothesis you are testing, and the approach you have taken to test it.

### So what do we put into the introduction?

As I’ve said elsewhere, the [hypothesis](hypothesis.html) itself is made up of different parts, and each of these must be explained in the introduction. We also need to understand the approach that you’ve decided to take in your study (e.g. experimental, lab or field approach, observations or natural history). All of these decisions that you made were informed by the literature, as is your general understanding of the subject that you are studying. So making sure that you cite ([read about citing](citations.html)) the relevant literature is a key ingredient of the introduction.

@@

### And the construction…?

Previously, I’ve described the introduction like a funnel (Figure \@ref(fig:introduction)), where we channel the reader into our hypothesis by starting broad and ending up narrow ([see the formula](formula.html)). This can draw criticism that journal articles are expected to fit a certain style, that this style is boring, and that what we really need is for people to write in a more exciting and varied way. Although I do not disagree with this, my objective here, is to try to demystify writing and enable you to get started. The easiest way to get started is, I believe, through a formula. But it is important to say that the funnel isn’t the only way, and I’ve read some great papers where the first sentence of the introduction is the hypothesis. However you do it, the hypothesis is at the heart of your introduction because it is the reason for your work.

@@

Table: (\#tab:introparagraphs) **Here is a guide to the contents of each paragraph of your introduction.** Although you should not feel constrained by this suggestion, it might help you get started when planning your bullet pointed structure of the introduction before you start writing.

@@

Remember that while this section on writing the introduction is short, much of [Part II](part2.html) concerns writing style that is applicable to your introduction. In particular, sections on composing your [hypothesis](hypothesis.html), writing a [paragraph](paragraph.html), and an [argument](argument.html) should be referred to when you are writing your introduction.

>>>>

# The Introduction {#introduction}

## Why do we need an introduction?
The introduction and discussion appear to be two common stumbling points for students writing chapters or manuscripts. First, what to put in and what to leave out. And second how to construct it. They are complementary, and if written well can be read without the Materials and Methods or Results, with the reader still gaining all of the major information of the manuscript.

The introduction is going to be the first part of your manuscript that anyone reads. Yes, they’ve already taken in the title and the abstract. These are almost like bait to draw readers in. The meat starts with the introduction. And it’s not only important to get all the correct content in there, but it’s also important not to give misleading information that might distract the reader.

When you write the introduction to a paper, you must not be tempted to stray away from what we are aiming to introduce. The aim of our introduction is to explain to the reader the hypothesis you are testing, and the approach you have taken to test it.

## So what do we put into the introduction?

The [hypothesis](#hypothesis) itself is made up of different parts, and each of these must be explained in the introduction. We also need to understand the approach that you’ve decided to take in your study (e.g. experimental, lab or field approach, observations or natural history). All of these decisions that you made were informed by the literature, as is your general understanding of the subject that you are studying. So making sure that you cite ([read about citing](#citations)) the relevant literature is a key ingredient of the introduction.

@@

## And the construction…?

Previously, I’ve described the introduction like a funnel (Figure \@ref(fig:introduction)), where we channel the reader into our hypothesis by starting broad and ending up narrow ([see the formula](#formula)). This can draw criticism that journal articles are expected to fit a certain style, that this style is boring, and that what we really need is for people to write in a more exciting and varied way. Although I do not disagree with this, my objective here, is to try to demystify writing and enable you to get started. The easiest way to get started is, I believe, through a formula. But it is important to say that the funnel isn’t the only way, and I’ve read some great papers where the first sentence of the introduction is the hypothesis. However you do it, the hypothesis is at the heart of your introduction because it is the reason for your work.

@@

Table: (\#tab:introparagraphs) **Here is a guide to the contents of each paragraph of your introduction.** Although you should not feel constrained by this suggestion, it might help you get started when planning your bullet-pointed structure of the introduction before you start writing.

@@

Remember that while this section on writing the introduction is short, much of [Part II](#part2) concerns writing style that is applicable to your introduction. In particular, sections on composing your [hypothesis](#hypothesis), writing a [paragraph](#paragraph), and an [argument](#argument) should be referred to when you are writing your introduction.
